---
title: "Joe's Useful Functions"
output: html_notebook
---

# Intro

This notebook serves as a compilation of miscellaneous functions I have found along my journey through R.

# Plot Functions

## Aesthetics

### Color Ramp Palette

You can create a custom color ramp palette function using the `colorramppalette` function stored as an object.

package needed: `RColorBrewer`

```{r}
#define a set of colors
colors <- RColorBrewer::brewer.pal(n = 8, name = "Dark2")

#make a color range using colorRampPalette()
#the object now becomes a function which to call a number on
color_range <- colorRampPalette(colors)

plot(rnorm(10),col = color_range(9), pch = 15)
```

## Plots

### mosaic plot

This function can be used to make a mosaic plot in `ggplot2`. The inputs are the data, X = the variable you are interested in seeing frequencies for, and the FILL is the variable you want to color by. 

packages needed: `ggplot2`, `reshape2`, `dplyr`, `ggthemes`
```{r, echo=TRUE}
library(ggplot2)
library(reshape2)
library(dplyr)
library(ggthemes)

# Script generalized into a function
mosaicGG <- function(data, X, FILL) {
  
  # Proportions in raw data
  DF <- as.data.frame.matrix(table(data[[X]], data[[FILL]]))
  DF$groupSum <- rowSums(DF)
  DF$xmax <- cumsum(DF$groupSum)
  DF$xmin <- DF$xmax - DF$groupSum
  DF$X <- row.names(DF)
  DF$groupSum <- NULL
  DF_melted <- melt(DF, id = c("X", "xmin", "xmax"), variable.name = "FILL")
  library(dplyr)
  DF_melted <- DF_melted %>%
    group_by(X) %>%
    mutate(ymax = cumsum(value/sum(value)),
           ymin = ymax - value/sum(value))
  
  # Chi-sq test
  results <- chisq.test(table(data[[FILL]], data[[X]])) # fill and then x
  resid <- melt(results$residuals)
  names(resid) <- c("FILL", "X", "residual")
  
  # Merge data
  DF_all <- merge(DF_melted, resid)
  
  # Positions for labels
  DF_all$xtext <- DF_all$xmin + (DF_all$xmax - DF_all$xmin)/2
  index <- DF_all$xmax == max(DF_all$xmax)
  DF_all$ytext <- DF_all$ymin[index] + (DF_all$ymax[index] - DF_all$ymin[index])/2
  
  # plot:
  g <- ggplot(DF_all, aes(ymin = ymin,  ymax = ymax, xmin = xmin,
                          xmax = xmax, fill = residual)) +
    geom_rect(col = "white") +
    geom_text(aes(x = xtext, label = X),
              y = 1, size = 3, angle = 90, hjust = 1, show.legend = FALSE) +
    geom_text(aes(x = max(xmax),  y = ytext, label = FILL),
              size = 3, hjust = 1, show.legend = FALSE) +
    scale_fill_gradient2("Residuals") +
    scale_x_continuous("Individuals", expand = c(0,0)) +
    scale_y_continuous("Proportion", expand = c(0,0)) +
    theme_tufte() +
    theme(legend.position = "bottom")
  print(g)
}

mosaicGG(diamonds, "clarity", "cut")
```



# Stat Functions

### Stat Quantile

Plot the quantiles of a range of a variable in a dataset using `stat_quantile` geom in `ggplot2`.

You can specify the quantile to plot as a value 0-1 where 0.25 is the first quantile, 0.5 = median, 0.75 = 3rd quantile, and 0,1 are the min and max, respectively.

```{r}
ChickWeight %>%
ggplot(aes(x = Time, y = weight, color = Diet))+
  geom_point(color = 'black')+
  stat_smooth(method = "lm", se = F)+
  stat_quantile(quantiles = 0.5, linetype = 2)

ChickWeight %>%
ggplot(aes(x = Time, y = weight, color = Diet))+
  geom_point(color = 'black')+
  stat_smooth(method = "lm")+
  geom_quantile(stat = 'quantile', quantiles = c(0, 1), linetype = 2)

```

